Software architectures to integrate workflow engines in science gateways

### 2016-05-14 paper.tex

In the last decade, the e-Science community has developed workflow
systems to help application developers access distributed
infrastructures such as clusters, grids, clouds and web
services. These efforts resulted in tools among which
Askalon~\cite{fahringer2005askalon},
Hyperflow~\cite{balis2016hyperflow}, MOTEUR~\cite{GLAT-08i},
Pegasus~\cite{deelman2005pegasus,Deelman201517},
Swift~\cite{zhao2007swift}, Taverna~\cite{oinn2004taverna},
Triana~\cite{taylor2007triana}, VisTrails~\cite{callahan2006managing}
and WS-PGRADE~\cite{Kacsuk2012} \note{vladimir}{As a past developer of it, I would suggest to refer to WS-VLAM as well, despite it is not very famous, if you do not mind :)}. Such workflow engines usually
describe applications in a high-level language with specific data and
control flow constructs, parallelization operators, visual edition
tools, links with domain-specific application repositories, provenance
recording and other features. An overview of workflow system
capabilities is available in~\cite{deelman2009workflows}.

@@

\paragraph{Robustness} No component is specific to workflow execution
(\texttt{R$_1$}=0), and only interaction \texttt{e} is necessary
(\texttt{R$_2$}=1).
\note{vladimir}{there is a large piece of text in Scalability about reliability of the workflow engine task which on my opinion suits Robustness better}

@@

Differences in other specific features should not be neglected. Nested
workflows (with service invocation) and workflow import (with service
invocation) are the only architectures that support meta-workflows. Besides,
task encapsulation and nested workflows may hamper
fine-grained debugging. 
The availability of fine-grained debugging information may also be
critical to efficient user support.

@@

Olabarriaga et al.~\cite{olabarriaga2014} presents a user-centered
view of the ecosystem of science gateways with focus on workflow
management. Their proposed layered system architecture includes
gateway, workflow engine and distributed infrastructure components,
similarly to the "service invocation" pattern. However other patterns
are not identified.

>>>>

In the last decade, the e-Science community has developed workflow
systems to help application developers access distributed
infrastructures such as clusters, grids, clouds and web
services. These efforts resulted in tools among which
Askalon~\cite{fahringer2005askalon},
Hyperflow~\cite{balis2016hyperflow}, MOTEUR~\cite{GLAT-08i},
Pegasus~\cite{deelman2005pegasus,Deelman201517},
Swift~\cite{zhao2007swift}, Taverna~\cite{oinn2004taverna},
Triana~\cite{taylor2007triana}, VisTrails~\cite{callahan2006managing}
and WS-PGRADE~\cite{Kacsuk2012} \note{vladimir}{As a past developer of it, I would suggest to refer to WS-VLAM~\cite{wsvlam} as well, despite it is not very famous, but it played an important role for Amsterdam e-science workflows}. Such workflow engines usually
describe applications in a high-level language with specific data and
control flow constructs, parallelization operators, visual edition
tools, links with domain-specific application repositories, provenance
recording and other features. An overview of workflow system
capabilities is available in~\cite{deelman2009workflows}.

@@

\paragraph{Robustness} No component is specific to workflow execution
(\texttt{R$_1$}=0), and only interaction \texttt{e} is necessary
(\texttt{R$_2$}=1).
\note{vladimir}{there is a large piece of text in Scalability about reliability of the workflow engine task which on my opinion suits Robustness better}
\note{vladimir}{in addition, the claim of very good robustness of task encapsulation seems questionable to me as in this model you delegate the critical task of controlling workflow execution to, let's say, some 'random' node on the grid that you have no control of, not sure about its performance, connectivity and reliability and so on; however, considering the models and dependencies presented in fig.2 this evaluation of robustness metric may be correct}

@@

Differences in other specific features should not be neglected. Nested
workflows (with service invocation) and workflow import (with service
invocation) are the only architectures that support meta-workflows. \note{vladimir}{As mentioned earlier by Silvia, this might not be exactly the case, depending on the point of view} Besides,
task encapsulation and nested workflows may hamper
fine-grained debugging. 
The availability of fine-grained debugging information may also be
critical to efficient user support.

@@

Olabarriaga et al.~\cite{olabarriaga2014} presents a user-centered
view of the ecosystem of science gateways with focus on workflow
management. Their proposed layered system architecture includes
gateway, workflow engine and distributed infrastructure components,
similarly to the "service invocation" pattern. However other patterns
are not identified.
\note{vladimir}{@silvia, does it make sense to mention workflow classification efforts based on workflow motifs and so on?}
